---
title: "How similar (or not) are WFMU DJs?"
output: html_notebook
author: Arthur Steinmetz
---
 

```{r}


library(devtools)
#install_github('sinhrks/ggfortify')
library(ggplot2)
library(stringr)
library(tm)
library(dplyr)
library(tidyverse)
library(lubridate)
library("ggplot2")
#library(ggfortify)
library("wordcloud")
library("RColorBrewer")
library("SnowballC")
library(igraph)
library(cluster)
# library(reshape2)
#library(proxy)
library(Matrix)

load("playlists.Rdata")
load("djkey.Rdata")

DJKey<-as_tibble(DJKey) %>% mutate(DJ=as.factor(DJ))
```

Clean up inconsistant artist names
```{r}

allDJArtists<-playlists %>% distinct() %>% filter(Artist != Title)

allDJArtists$ArtistToken<-allDJArtists$Artist
# one artist is all punctuation so give !!! special treatment
allDJArtists$ArtistToken<-str_replace(allDJArtists$ArtistToken,"!!!","chkchkchk")
# now change some common punctuation to space
print("Stripping Punctuation")

allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"\\&"," ")

allDJArtists$ArtistToken<-str_to_lower(allDJArtists$ArtistToken)
# I choose to strip out the stuff below though dealing with it might get better analysis
#remove any text in parentheses
print("Stripping filler words")
# get rid of anything between parenthesis
#tricky regex to handle cases of multiple parentheticals in one artist
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"(\\([^(]+\\))","")
# remove 'featuring' or 'with' artists
# I chose not to remove "Versus" because that is a band name
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"(feat |featuring |and the |with |vs |vs\\.).+","")
# get rid of 'live' identifier
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"(live @ |live on|@).+","")

#now get rid of remaining non-word characters except space

allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"[^a-z^ ^0-9]","")


# get rid of 'interview'
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"(interview w|interview)","")

# get rid of unspecified artists
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"unknown artist(s| )|unknown","")
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"various artists|various","")

#get rid of the marathon finale
allDJArtists<-allDJArtists%>%filter(!str_detect(Artist,"hoof[a-zA-Z ]+sinfonia"))

#make "new york" one word.  Lots of bands start with the term
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"new york","newyork")

#make "x ray" one word. hopefully we've stripped out the dash already.Lots of bands start with the term
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"x ray","xray")

#now some connecting words that might be spelled/used variantly
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"and | of | the "," ")

#and leading "the"
allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"^the "," ")
# strip leading/trailing whitespace
allDJArtists$ArtistToken<-str_trim(allDJArtists$ArtistToken)

#did we create any null entries
allDJArtists<-filter(allDJArtists,Artist!="")
allDJArtists<-filter(allDJArtists,Artist!="Artist")

allDJArtists<- filter(allDJArtists,!str_detect(Artist, "Music Behind Dj:"))
allDJArtists<- filter(allDJArtists,!str_detect(Artist, "Wake N Bake"))


numWords=2 #is two enought for uniqueness?

  # we replaced all punctuation with spaces
  #maybe strip spaces and combine all artist Words
  #combine first two words
  print("Trying to make sense of artist names")
  #does this break if numWords> number of words?
  t<-str_split_fixed(allDJArtists$ArtistToken,pattern="[ ]+",n=numWords+1)[,1:numWords]
  
  allDJArtists$ArtistToken<-apply(t,MARGIN=1,FUN=paste,collapse="")

  # There are a dozen ways Andy Breckman can misspell "Bruce Springsteen."
  allDJArtists<- allDJArtists %>%   
    mutate(ArtistToken=replace(ArtistToken,
                               str_detect(ArtistToken,'brucesp'),
                               "springsteen"))
  
  #Code below used for unique list of artists. Not used here.
  #now that tokens are created extract unique ones for each dj so mulitples don't occur
  # the zillion flavors of "Sun Ra..." will show up for each DJ only once
  # not perfect.  There are a dozen ways Andy Breckman can misspell "Bruce Springsteen."
  #print("Create list of unique artist names for each DJ")
  #artistTokens<-allDJArtists%>%select(DJ,artistToken)%>%group_by(DJ)%>%distinct(artistToken)


  print("Combining iconic 2-name artists into one name to save space in wordcloud")
  allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"rollingstones","stones")
  allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"enniomorricone","morricone") #only on WFMU!
  allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"davidbowie","bowie")
  allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"bobdylan","dylan")
  allDJArtists$ArtistToken<-str_replace_all(allDJArtists$ArtistToken,"yola","yolatengo")
  
  allDJArtists<-allDJArtists %>% 
    mutate(DJ=as.factor(DJ)) %>%
    filter(ArtistToken !="") %>% 
    filter(ArtistToken !="yourdj") %>% 
    filter(ArtistToken !="hoofmouth") %>% 
    filter(ArtistToken !="tomwilson") %>%  #not songs
    distinct() %>% #why would there be dupes?  Don't know, but there are
    group_by(DJ) 
```
Data is reasonably clean.  Do some analysis
```{r}

#playlists by year
allDJArtists %>% 
  ungroup() %>% 
  mutate(Year=trunc(decimal_date(AirDate)))  %>% 
  group_by(Year) %>% 
  summarise(Playlists=n()) %>% 
  arrange(Year) %>% 
  filter(YrQtr> as.yearqtr("2000-01-01")) %>% 
  ggplot(aes(x=Year,y=Playlists))+geom_col()



```

```{r}

#-------------------------------------------------------------------------------------
# We love wordclouds!
#scalefactor magnifies differences for wordcloud
#use filters to select on or off schedule and constrain show airdate range
scaleFactor=2
maxWords = 200
  #most popular artists
  top_artists<-DJKey %>% 
    #filter(onSched==TRUE) %>% #on Sched or off?
    left_join(allDJArtists)%>%
    group_by(ArtistToken)%>%
    filter(AirDate<as.Date("2017-03-18")) %>%  #date range?
    summarize(play_count=n())%>%
    arrange(desc(play_count))
wordcloud(words = top_artists$ArtistToken, freq = top_artists$play_count^scaleFactor,max.words=maxWords, random.order=FALSE,rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"),scale = c(3,.3))

```
#Dead Artist Effect. When did interest in an artist increase?
```{r}
dead_artists=c('bowie','prince','amywinehouse','sharonjones','chuckberry')

allDJArtists %>% 
  ungroup() %>% 
  filter(ArtistToken %in% dead_artists) %>% 
  mutate(YrQtr=as.yearqtr(AirDate))  %>% 
  group_by(ArtistToken,YrQtr) %>% 
  summarise(Spins=n()) %>% 
  arrange(YrQtr) %>% 
  filter(YrQtr> as.yearqtr("2000-01-01")) %>% 
  ggplot(aes(x=YrQtr,y=Spins))+geom_col()+facet_wrap(~ArtistToken)

```

combine artist list into a single string per DJ
```{r}
#-------------------------------------------------------------  
#combineAllArtists
  concat_artists<- data_frame()
  #make sure there aren't extra levels
  allDJArtists$DJ<-factor(allDJArtists$DJ,as.character(unique(allDJArtists$DJ)))
  for (dj in levels(allDJArtists$DJ)){
    #put all words in string for each DJ
    print(paste("Creating artist documents",dj))
  concat_artists<-bind_rows(concat_artists,data_frame(DJ=dj,
                        Artists= allDJArtists%>%
                          filter(DJ==dj)%>%
                          pull(ArtistToken)%>% 
                          paste(collapse=" ")%>%
                          str_replace_all("[^a-z ]","")%>%as.character(),
                        onSched=DJKey%>%filter(DJ==dj)%>%pull(onSched)))
  }
  #artists should not have factor levels
  #concat_artists$Artists<-as.character(concat_artists$Artists)
  concat_artists<-filter(concat_artists,Artists!="")
```
Create term document matrices
```{r}
print("Create document corpus and term document matrices")
djCorpus <- VCorpus(VectorSource(concat_artists$Artists))


for (i in 1:length(djCorpus)) {
  meta(djCorpus[[i]], tag="id") <- concat_artists$DJ[i]  
  meta(djCorpus[[i]], tag="DJ") <- concat_artists$DJ[i]
  meta(djCorpus[[i]], tag="onSched") <- (DJKey %>% 
    filter(DJ==concat_artists[1,1] %>% 
             pull()) %>% 
    pull(onSched))
}



sched_status=c("on","off","all")
status<-sched_status[1]
#idx <- switch(mic,
#              on = (meta(djCorpus, "onMic") == TRUE),
#              off = (meta(djCorpus, "onMic") == FALSE),
#              both = rep(TRUE,length(djCorpus))
#              )


#djtdm<-TermDocumentMatrix(djCorpus[idx])%>%removeSparseTerms(SPARSE)

#OR if you have the memory to pre-create
#make 3 Term Document Matrices where artist is in the column based on onmic status
#get roughly top 400 artists when removeSparseTerms(0.80) used. top 8000 when 0.95 sparse is used
SPARSE<- 0.95 #sparsity of term document matrices

djtdm_all<-TermDocumentMatrix(djCorpus) #%>%removeSparseTerms(SPARSE)
djtdm_on<-TermDocumentMatrix(djCorpus[meta(djCorpus, "onSched") == TRUE])%>%removeSparseTerms(SPARSE)
djtdm_off<-TermDocumentMatrix(djCorpus[meta(djCorpus, "onSched") == FALSE])%>%removeSparseTerms(SPARSE)

# now create Document Term Matrix where DJs are the column
djdtm<-DocumentTermMatrix(djCorpus) #%>%removeSparseTerms(SPARSE)
```

#Analyze similarity
```{r}
jaccard <- function(m) {
  #http://stackoverflow.com/questions/36220585/efficient-jaccard-similarity-documenttermmatrix
  ## common values:
  A <- tcrossprod(m)
  im <- which(A > 0, arr.ind=TRUE, useNames = FALSE)
  b <- rowSums(m)
  Aim <- A[im]
  J<-sparseMatrix(
    i = im[,1],
    j = im[,2],
    x = Aim / (b[im[,1]] + b[im[,2]] - Aim),
    dims = dim(A)
  )
  
  #preserve row/column names, if any
  rownames(J)<-rownames(m)
  colnames(J)<-rownames(m)
  return( J )
}  
#-----------------------------------------------------------
getSimilarity<-function(djdtm=djdtm){
  m2<-as.matrix(djdtm)
  rownames(m2)<-djDocs$DJ

  # get similarity
  #j<-simil(m2,method="Jaccard")
  j<-jaccard(m2)
  save(j,file="djSimilarity.RData")
  return(j)
}
#-----------------------------------------------------------


# get similarity index matrix using Jaccard
j<-getSimilarity(djdtm)

#cluster plot of DJs
set.seed(1)
CLUSTERS=5
  kdj<-kmeans(j,CLUSTERS)
  clust<-kdj$cluster
  #djCluster<-cbind(DJ=names(clust),data.frame(cluster=clust))
  
  #if ("cluster" %in% names(DJKey)){
  #  # then remove the cluster column
  #  DJKey<-select(DJKey,-cluster)
  #}
  # now replace with new clustering
  #DJKey<-inner_join(DJKey,djCluster)%>%arrange(cluster)
    
  
  clusplot(as.matrix(j), clust, main="DJ Similiarity Clusters",color=T, shade=T, labels=2, lines=0) 
}
```

```{r}
#----------------------------------------------------------
similarDJs<-function(whichDJ="TW",compareDJ="TM"){
  #djs for example
  #whichDJ <- "TW"
  #compareDJ<-"TM"
likeDJs<-data_frame(DJ=row.names(j),similarity=as.matrix(j)[,whichDJ]) %>% 
  arrange(desc(similarity)) #sort descending and remove self DJ

  #since we sorted in order of descending similarity the first row has the most similar DJ
  compareDJ<-likeDJs[1:3,1] %>% pull() 
  whichShow<-DJKey%>%filter(DJ==whichDJ)%>%.$ShowName%>%as.character()
  compareShow<-DJKey%>%filter(DJ %in% compareDJ)%>%.$ShowName%>%as.character()
  print(paste(whichShow,"is most similar to",compareShow))
  print(paste("Similarity Index:",format(pull(likeDJs[1:3,2]),digits=2),"/1.00"))
  
  commonArtists<-intersect(artistTokens[which(artistTokens$DJ==whichDJ),]$artistToken,artistTokens[which(artistTokens$DJ==compareDJ),]$artistToken)
  print(data.frame(Common_Artists=sample(commonArtists,20)))
}

similarDJs()
```

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
