---
title: "Rock Deaths 2"
output: html_notebook
---

#Bonus analysis.  Where did they die?
```{r}
deaths3<-deaths %>% 
  separate(Location,c("neighborhood","city","state","country"),
           remove=FALSE,
           sep=", ",
           fill="left",
           extra="merge")

deaths3<-deaths3 %>% mutate_cond(country=="",country="Not Specified")
deaths3<-deaths3 %>% mutate_cond(is.na(city),city="Not Specified")
deaths3<-deaths3 %>% mutate_cond(country=="U.S.",country="USA")
deaths3<-deaths3 %>% mutate_cond(country=="U.K.",country="England")
deaths3 %>% group_by(country) %>% 
  summarise(count=n()) %>% 
  arrange(desc(count)) %>%
  .[1:12,] %>% 
  kable(format.args=list(big.mark=","))

```

This doesn't reveal any surprises.  Again, it says something about the home of rock and the location of people who submit entries to Wikipedia.  We could go further and compare these numbers to population but we wouldn't learn anything not already obvious.

What about cities?

```{r}
deaths3 %>% group_by(city) %>% 
  summarise(count=n()) %>% 
  arrange(desc(count))%>% 
  .[1:12,] %>% 
  kable(format.args=list(big.mark=","))
```
If we looked at city populations, we'd find that L.A., Nashville and Memphis have more rock musicians than population would suggest. Again, not a surprise.  It's a bit surprising to me that London is not on the list. On second thought, every English rock star I can think of moved to the U.S.  I occasionally walk by the flowers outside David Bowie's house in Manhattan.
