---
title: "R Notebook"
output: html_notebook
---

Setup environment
```{r}
# scrape playlists

library(rvest)
library(stringr)
library(xml2)
library(tidyverse)

# ----------------------------------------------
ROOT_URL<-"http://wfmu.org"
#limit analysis to DJs with at least numShows shows.
# This also excludes DJs where we couldn't extract valid playlist URLs.
numShows <- 50

```
Get URLS of the the web pages containin links to the playlists
```{r}

#-------------------------------------------
getDJURLs <- function(){
  rawDJURLs<- read_html(paste(ROOT_URL,"/playlists",sep=""))
  # get the urls of the each DJs RSS playlist feed
  t<-rawDJURLs%>%html_nodes(xpath='//html//body//center[2]//table[1]//table//a[contains(.,"Playlists")]') %>% 
    html_attr(name="href")

  DJURLs<-paste("http://wfmu.org",t,sep="")[-1]
  # above got the RSS feed links but we want the longer list of shows.  Below modifies
  # the URL to get the right link
  DJURLs<- gsub("playlistfeed","playlists",DJURLs)
  DJURLs<- gsub(".xml","",DJURLs)
  
  return(DJURLs)
}
DJURLs<-getDJURLs()
```
If the DJKey file is already built load it.  If not run the chunk following the loading step to build it.
```{r}
#load(file='djkey.rdata')

```

Get the show names for each DJ and identify those that are not currently on the air.  Here is where we build the DJKey data frame initially.  If the file is already built, load it instead of running the chunk.
```{r}
#-------------------------------------------
getDJsOffSched <- function(){
  #table 9 is off sched. 2-8 are monday through sunday
  rawDJURLs<- read_html(paste(ROOT_URL,"/playlists",sep=""))
  t_off<-rawDJURLs%>%html_nodes(xpath='//html//body//center[2]//table[1]//table[9]//a[contains(.,"Playlists")]')  %>% html_attr(name="href") 
  d_off <- str_replace(t_off,"/playlistfeed/","")
  d_off <- str_replace(d_off,".xml","")
  return(d_off)
}

#---------------------------------------------------
# get the shownames for a DJ
getShowNames<-function(DJURLs) {
  DJKey <- data.frame()
  for (page in DJURLs) {
    singleDJ<- read_html(page)
    showName <- html_node(singleDJ,"title")%>%html_text()
    showName <- gsub("\n","",sub("Playlists and Archives for ","",showName))
    showName<-str_replace(showName,'WFMU:',"")
    showName<-str_replace_all(showName,':Playlists and Archives',"")
    DJ <- sub("http://wfmu.org/playlists/","",page)
    DJKey<-rbind(DJKey,data.frame(DJ=DJ,ShowName=showName))
    print(showName)
  }
  # now identifty those DJs which are currently on the schedule
  DJKey$onSched <- 'YES'
  DJKey$onSched[which(DJKey$DJ %in% getDJsOffSched())]<-'NO'
  #strip "WFMU" and "Playlists and Archives" and some punctuation
  DJKey$ShowName<-str_replace_all(DJKey$ShowName,"(P|p)laylists (and|&) (A|a)rchives","")
  DJKey$ShowName<-str_replace_all(DJKey$ShowName,"-","")
  DJKey$ShowName<-str_replace_all(DJKey$ShowName,"(P|p)laylist|(R|r)ecent","")
  DJKey$ShowName<-str_replace_all(DJKey$ShowName,"WFMU|wfmu","")
  DJKey$ShowName<-str_replace_all(DJKey$ShowName,"The ","")
  DJKey$ShowName<-str_trim(DJKey$ShowName)
  

  return (DJKey)  
}
DJKey<-getShowNames(DJURLs)
save(DJKey,file="DJKey.RData")

```
Eliminate non-music DJ's from the DJKey table.  This list is created by hand and was last updated at the end of 2017.
```{r}
excludeDJs <-
  c('SD',
    'HA',
    'BC',
    'AF',
    'CP',
    'HP',
    'JP',
    'GM',
    'DC',
    'CC',
    'DU',
    'ES',
    'LW',
    'IM',
    'LL',
    'NW',
    'NP',
    'ZZ',
    'FC',
    'SY',
    'TI',
    'LK',
    'TP',
    'RC',
    'VC')
music_djs<-DJKey %>% 
  select(DJ) %>% 
  anti_join(data_frame(DJ=excludeDJs)) %>% 
  pull(DJ)
```
Get the URLs of the pages actually containing the playlists.  The tricky bit here is that some DJ pages contain all their playlists going back multiple years and some pages contain just the current year with further links to prior years.  I try to catch this and go into the next level down.

```{r}
# -------------get the URLs of the playlist pages for a DJ ----------
#should work to delve into earlier years
get_playlist_page_URLs<-function(url_suffix) {
  #first call should be the base DJ page with links to any earlier year playlist lists
  if (str_length(url_suffix)==2){
    dj<-url_suffix
    latest_url<-paste0("/playlists/",url_suffix)
    url_suffix<-latest_url
  }
  singleDJ<- read_html(paste0("http://wfmu.org",url_suffix))
  #this assumes the earlier year playlist links are of the form
  # wfmu.org/playlists/<dj><year>/
  pl_url<-singleDJ %>%
    html_nodes(xpath=paste0("//a[contains(@href,'playlists/",dj,"')]")) %>%
    html_attr("href")
  pl_url<-c(latest_url,pl_url) %>% unique()
  
return(pl_url)
}

# #---------------------------------------------------
# get the URLs of the playlists for a DJ
getDJPlaylistURLs<-function(music_djs) {
  DJ_playlists = NULL
  dudList<-NULL
  #DJKey = data.frame()
  for (dj in music_djs) {
    print(dj)
    url_suffixes<-get_playlist_page_URLs(dj)
    for (u in url_suffixes){
      print(u)
      singleDJ<- read_html(paste0("http://wfmu.org",u))
      pl<-singleDJ%>%
        html_nodes(xpath="//a[contains(@href,'playlists/shows')]") %>%
        html_attr("href")
      #format for newer shows
      pl<-as.character(na.omit(pl[str_detect(pl,"playlists/shows")]))
      # format for older shows
      if (length(pl)<1) pl<-as.character(na.omit(pl[str_detect(pl,"Playlist")]))
      
      #assume a full URL is a fill-in DJ.  We omit these from the analysis
      pl<-pl[!str_detect(pl,"http")]
      
      playlistURL<-pl %>% as.character()
      #omit shows without valid playlists.  Talk shows?
      if (length(playlistURL)>0) {
        DJ_playlists = bind_rows(DJ_playlists, data_frame(DJ=dj,playlistURL = playlistURL))
      } else { 
        print("DUD")
        dudList<-c(dudList,dj) }
      
    }
  }  
  
  return(DJ_playlists)
}

playlistURLs<-getDJPlaylistURLs(music_djs)

```
Add show counts to the DJKey table.
```{r}
showCounts<-playlistURLs %>% 
  group_by(DJ) %>% 
  summarise(showCount=n()) %>% 
  arrange(desc(showCount))
DJKey<-left_join(DJKey,showCounts)
# non-music shows
djList <- DJKey %>% 
  filter(showCount > numShows, !(DJ %in% excludeDJs)) %>%
  pull(DJ)

```

The chunk below is essentially depricated.  Instead of gettting this actual playlists, this function gets the lists of every artist the DJ has played, which is displayed on each DJ's playlist page. I keep the function here but it is not called anywhere.

```{r}
#-------------------------------------------------
# Get all Artists ever played by a DJ
#WFMU maintains this as a separate page
getDJArtistNames<-function(DJURLs) {
  # scrape artist names for all DJs from the link at the bottom of each DJ page
  allDJArtists<-data.frame()
  URL_BRANCH<- "/artistkeywords.php/"
  for (page in DJURLs) {
    singleDJ<- read_html(page)
    showName <- html_node(singleDJ,"title")%>%html_text()
    showName <- gsub("\n","",sub("Playlists and Archives for ","",showName))
    DJ <- sub("http://wfmu.org/playlists/","",page)
    DJKey<-rbind(DJKey,data.frame(DJ=DJ,ShowName=showName))
    print(showName)
    artistListPage <- paste(ROOT_URL,URL_BRANCH,DJ, sep="")
    artistList<-read_html(artistListPage)%>%html_node(xpath="//body/div")%>%html_text()%>%str_split("\n")
    DJArtists<-data.frame(DJ,artistRaw=unlist(artistList))
    if (nrow(DJArtists) >0) allDJArtists = rbind(allDJArtists,DJArtists)
    #remove factor level of DJs with no artists
    save(allDJArtists,file="allDJArtists.RData")
  }
  return(allDJArtists)
}  
```

```{r}

#--------------------------------------------------------------------------
altArtistNames <- c('THE STOOGE', 'Band', 'Singer', 'Artist')
altTitleNames <- c('THE SONG', 'Track', 'Song','Title')
altHeaderNames <- c(altArtistNames, altTitleNames)
MAXLEN = 60 #how long should we let artist and title names be. Truncate longer
header_th_xpath <- paste(
  "//th='Track'",
  "or //th='Title'",
  "or //th='Song'",
  "or //th='THE SONG'"
)
header_td_xpath <- paste(
  "//td='Track'",
  "or //td='Title'",
  "or //td='Song'",
  "or //td='THE SONG'",
  "or //td='Artist'",
  "or //td='THE SONG'",
  "or //td='THE STOOGE'"
)

```

The code below constains DJ-specific scraping routines.  They get called when other methods fail.  I could choose to call them just when that specific DJ comes up but I call them on all fails on the off chance that another DJ has the same playlist format.
```{r}
#------------------------------------------------------------------
try_BK<-function(wp){
  #Bill Kelly's format
  #assume first field is title and second is artist separated by dash
  #doing it it two steps assure same number of artists and titles
  title_artist<-wp %>% 
    html_nodes(xpath="//table[2]") %>% 
    html_text() %>% 
    str_extract_all("\n[\\S ]+\n-\n[\\S ]+\n") %>% 
    .[[1]] 
  
  
  Title<-title_artist %>% 
    str_extract_all("\n[\\S ]+\n-") %>% 
    str_replace_all("\n(-)?","")
  
  Artist<-title_artist %>% 
    str_extract_all("\n-\n[\\S ]+\n") %>% 
    str_replace_all("\n(-)?","")
  plraw<-tibble(Artist,Title)
  
}
#------------------------------------------------------------------
try_HN<-function(wp){
  
  #assume first field is title and second is artist separated by colon
  #doing it it two steps assure same number of artists and titles
  title_artist<-wp %>% 
    html_nodes(xpath="//table[2]") %>% 
    html_text() %>% 
    str_extract_all("\n[\\S ]+: [\\S ]+\n") %>% 
    .[[1]] 
  
  
  Artist<-title_artist %>% 
    str_extract_all("\n[\\S ]+: ") %>% 
    str_replace_all("(\n)|(: )","")
  
  Title<-title_artist %>% 
    str_extract_all(": [\\S ]+\n") %>% 
    str_replace_all("(\n)|(: )","")
  plraw<-tibble(Artist,Title)
  
}#------------------------------------------------------------------

fixHeaders <- function(pl) {
  #takes a data frame
  nm<-which(names(pl) %in% altArtistNames)
  tt<-which(names(pl) %in% altTitleNames)
  if (length(nm)>0){
    names(pl)[nm] <- "Artist"
  } else{
    pl=NULL
    return(pl)
  } 
  if (length(tt)>0){
    names(pl)[tt] <- "Title"
  } else{
    pl$Title=NA
  }
  
  return(pl)
}

#--------------------------------------------------------------------
get_playlist <- function(plURL, dj) {
  
  wholepage <- read_html(paste0(ROOT_URL, plURL))
  #try to pull out the show date.  assume first date in text on page is the show date
  airDate <- wholepage %>%
    html_text() %>%
    str_extract('[A-Za-z]+ [0-9]{1,2}, [0-9]{4}') %>%
    as.Date("%B %d, %Y")
  if (is.na(airDate)) {
    #try something else
    airDate <- wholepage %>%
      html_text() %>%
      str_extract('[0-9]{1,2} [A-Za-z]+ [0-9]{4}') %>%
      as.Date("%d %B %Y")
    
  }
  print(airDate)
  if (airDate < most_recent_date-7) {
    print("Back Far Enough")
    return("stop")
  }
  plraw <- NULL
  #hand-rolled
  #simplest case. A table with obvious header names
  if (!is.na(wholepage %>% html_node(xpath = "//th[@class='song']"))) {
    table_shell<-xml_new_root("table")
    #remove single column rows, I hope nothing else.
    wholepage %>% html_nodes(xpath="//td[@colspan='8']") %>% xml_remove(free=T)
    plraw<-wholepage%>% 
      html_nodes(xpa="//tr[td[@class ='song']] | //tr[th[@class ='song']]")
    for (node in plraw)  xml_add_child(table_shell,node)
    plraw<-table_shell %>% 
      html_node(xpath="//table") %>% 
      html_table(fill=TRUE)
    
  } else {
    # no 'th' but are there rows in a table with td of class=song?  get the table
    if (!is.na(wholepage %>% html_node(xpath = "//td[@class='song']"))) {
      plraw <- wholepage %>%
        html_node(xpath = "//td[@class='song']/ancestor::table") %>%
        html_table(fill = T)
      #now find the row that has the header
      for (n in 1:nrow(plraw)) {
        if (TRUE %in% (plraw[n, ] %in% altHeaderNames)) {
          names(plraw) <- plraw[n, ]
          plraw <- plraw[n + 1:nrow(plraw), ]
          break
        }
        if (n == nrow(plraw)) {
          print("DUD. Can't find header")
          plraw <- NULL
        }
      }
    }
  }
  
  if (is.null(plraw)) {
    # no song class, now what? is it a table? try to  find header
    #seems like cellspacing means its a row column thing
    pl_table<-wholepage %>% html_node(xpath = "//table[@cellspacing and @cellpadding]")
    num_rows<-pl_table %>% html_nodes("tr") %>% length()
    if (num_rows>2) {
      pl_table<-html_table(pl_table,fill = TRUE)
      if (any(names(pl_table) %in% altHeaderNames)) {
        plraw<-pl_table
      } else{
        # try one more ``
        #scan until we find the playlist header
        for (n in 1:nrow(pl_table)) {
          if (TRUE %in% (pl_table[n,] %in% altHeaderNames)) {
            names(pl_table) <- pl_table[n, ]
            plraw <- pl_table[n + 1:nrow(pl_table), ]
            break
          }
        }
      }
    }
  }
      
# SPECIAL DJ TREATMENT
  if (is.null(plraw)){
    #try idiosyncratic djs
    if (dj=="TW") plraw<-try_BK(wholepage)
    if (dj=="HN") plraw<-try_HN(wholepage)
    if (is.null(plraw)) print("DUD. Can't find header")
    #final dead end
    }
    
  plraw<-fixHeaders(plraw)
  # final clean up if we have something
  if (is.null(plraw)) {
    playlist <- NULL
  } else {
    if (TRUE %in% is.na(names(plraw)))
      plraw <- plraw[, -which(is.na(names(plraw)))] #sometimes an NA column
    
    playlist <- plraw %>%
      select(Artist,Title) %>% 
      na.omit() %>% 
      transmute(DJ = dj, 
                AirDate = airDate,
                Artist= Artist %>% 
                  str_trunc(MAXLEN,"right") %>% 
                  str_to_title() %>% 
                  str_extract("[\\S ]+") %>%         #get rid of extra lines in mutiple line cells
                  str_extract("[^\\(]+") %>%         #dump parentheticals
                  str_trim(),
                Title =  Title %>% 
                  str_trunc(MAXLEN,"right") %>% 
                  str_to_title() %>%
                  str_extract("[\\S ]+[^()]") %>% 
                  str_extract("[^\\(]+") %>%
                  str_trim()
      ) %>% 
      filter(Artist != '') %>% 
      filter(!is.na(Artist))
    print(playlist[1:5, ])
    
  }
  
  return(playlist)
}

```

Main loop

I don't like this much.  It makes an assumption that DJs have just one show per week in determining how far back to go when updating and existing playlist file.  A better approach would be to go by each DJ.  Find the latest date for each dj in the existing file, extract the show number for that date, then reject all URLs with lower show numbers. We still have to start with every creating every URL. 
```{r include=FALSE}
#-------------- MAIN -----------------


showCounts<-playlistURLs %>% 
  group_by(DJ) %>% 
  summarise(showCount=n()) %>% 
  arrange(desc(showCount))
DJKey<-left_join(DJKey,showCounts)
#limit analysis to DJs with at least numShows shows.
# This also excludes DJs where we couldn't extract valid playlist URLs.
numShows <- 50
# non-music shows
djList <- DJKey %>% 
  filter(showCount > numShows, !(DJ %in% excludeDJs)) %>%
  pull(DJ)

#careful not to trash intermediate results!
#playlists_raw = data_frame()
load("~/R Projects/wfmu/playlists_raw.RData")
UPDATE_ONLY =TRUE
if (UPDATE_ONLY) {
  #assume at most 5 shows per week
  #most shows are 1/week except the morning show
  most_recent_date<-max(playlists_raw$AirDate)
  go_back_num<- as.integer(round ((Sys.Date() - most_recent_date) * 5/7))
  print(paste("Updating ONLY last",go_back_num,"Shows"))
} else{
  go_back_num<-as.Date("1900-01-01") #arbitrarily far back so it's not a binding constraint
}

# if this breaks at a certain dj, correct and start with that DJ, in this case "TW."
# line below will process all djs after the matched DJ
#ideally not needed
#djList<-djList[match("TW",djList):length(djList)]
for (dj in djList) {
  plURLs <- playlistURLs %>%
    filter(DJ == dj) %>%
    rowid_to_column() %>% 
    select(playlistURL)
  if (UPDATE_ONLY) {
    print(paste("Updating at most last",go_back_num,"Shows to",most_recent_date))
  } else{
    go_back_num<-nrow(plURLs)
  }
  for (n in 1:go_back_num){
    plURL<-plURLs[n,1]
    print(paste(dj, n, plURL))
    if (!is.na(pull(plURLs[n,1]))){
      new_playlist<- get_playlist(plURL, dj)
      #end loop if we have reached most_recent_date beacuse get_playlist returns "stop" character string.
      if (TRUE %in% (is(new_playlist)=="character")) {
        break
      }else{
        playlists_raw <- bind_rows(playlists_raw,new_playlist)
      }
    }
  }
  #save to disk after each dj
  if (!(TRUE %in% (is(new_playlist)=="character"))) {
    print(paste("Saving playlists at",Sys.time()))
    save(playlists_raw,file="playlists_raw.Rdata")
  }
}

bad_Tables<-anti_join(tibble(DJ=djList),playlists_raw)

playlists_raw<-playlists_raw %>% 
  filter(Artist != Title) %>% #single column span across table.  Not a song.
  distinct()

save(playlists_raw,file="playlists_raw.Rdata")

```
