---
title: "A Statistical Examination Playlists WFMU.org: Supplement 1."
output: 
  html_notebook: 
    code_folding: hide
author: Arthur Steinmetz
---
# Answering some of Ken's questions 

```{r, message=FALSE, warning=FALSE}


library(devtools)
library(ggplot2)
library(stringr)
library(tm)
library(dplyr)
library(tidyverse)
library(lubridate)
library(broom)
library(ggrepel)
library("ggplot2")
library(ggfortify)
library(cowplot)
library(xts)
library("wordcloud")
library("RColorBrewer")
#library("SnowballC")
library(igraph)
library(cluster)
# library(reshape2)
library(vegan) #similarity measures
library(Matrix)
library(knitr)

source('arcdiagram.r')
load("playlists.rdata")
load('djkey.rdata')
DJKey$DJ<-as.character(DJKey$DJ)
```
Ken: On the Top Ten by year and in a few other places I see things that are parts of theme songs or that kind of thing.. like Steve Martin "Under the Bamboo Tree" which is part of Frank OToole's into song/collage. 

Art: As I noted I tried to catch these but have undoubtedly missed some.  Where you see specific song's that distorts results you would like to see deleted, let me know.

Frank's a weird case but his intro song does introduce distortions.  Let's strip out ALL of Frank's first plays because the same song is listed so many different ways. He does play a lot of Beach Boys, intro and otherwise.
```{r}
#careful.  every time you run this is will strip out the first song from each airdate,
#creating a new first song.  Eventually, there will be nothing left.

frank<-playlists %>% filter(DJ=="FX") #only frank
playlists<-playlists %>% filter(DJ!="FX") #frank is gone
songs_to_strip<-frank %>% 
  group_by(AirDate) %>% 
  summarize(first_song=first(Title)) %>%
  group_by(first_song) %>% 
  summarize(count=n()) %>% 
  pull(first_song)

frank<-frank %>% filter(!(Title %in% songs_to_strip))
playlists<-playlists %>% bind_rows(frank) #frank is back
```
Ken: Could I see the DJ Similarity Clusters in a larger format so I can see what's going on better? It's hard to read the two letter codes and the band names. 

Art: Let's try tabular format for now.
```{r}

```

Ken: I would love to see a simple ranking of the most played songs. 
```{r}
top_songs <- playlists %>% 
  mutate(artist_song=paste(ArtistToken,Title))

top_100<- top_songs %>% 
  group_by(artist_song) %>% 
  summarise(play_count=n()) %>% 
  arrange(desc(play_count)) %>% 
  top_n(100)

top_100_last5<- top_songs %>% 
  filter(AirDate>Sys.Date()-(365*5)) %>% 
  group_by(artist_song) %>% 
  summarise(play_count=n()) %>% 
  arrange(desc(play_count)) %>% 
  top_n(100)

top_songs %>% 
  filter(artist_song=="Focus Hocus Pocus") %>% 
  group_by(DJ) %>% 
  summarise(count=n()) %>% 
  arrange(desc(count))
```




